\documentclass[../thesis.tex]{subfiles}

\begin{document}
\vspace{-1\baselineskip}

\section{Data samples}
\label{sec:data}
Data samples used in this analysis were collected by the \acs{ATLAS} detector during Run 2 data-taking campaign between 2015-2018. The samples contain $pp$ collisions at center-of-mass energy of $\sqrt{s}=13$ TeV with 25 ns bunch-spacing, which corresponds to an integrated luminosity of 140 \fb with an uncertainty of 0.83\% \citep{DAPR-2021-01}. The \acs{HLT} trigger strategy is similar to that of previous \tttt observation analysis \citep{tttt_obs} and include single lepton and dilepton triggers. Calibration for di-muon and electron-muon triggers were not ready for the samples used in this analysis, and are therefore not included. Events are also required to contain at least one lepton matched to the corresponding object firing the trigger. Triggers used are summarized in \autoref{tab:ana:trigger}.

\begin{table}[!ht]
\centering
\caption{\label{tab:ana:trigger}Summary of all \acs{HLT} triggers used in this analysis. Events are required to pass at least one trigger.}%
\begin{tabular}{p{9cm}|cccc}
\toprule\toprule
\multicolumn{1}{c|}{\multirow{ 2}{*}{Trigger}}	& \multicolumn{4}{c}{Data period} \\
\multicolumn{1}{c|}{}							& 2015	& 2016	& 2017	& 2018 \\
\midrule
\multicolumn{5}{c}{Single electron triggers} \\
\midrule
\verb|HLT_e24_lhmedium_L1EM20VH| 		& \checkmark & - & - & - \\
\verb|HLT_e60_lhmedium|					& \checkmark & - & - & - \\
\verb|HLT_e120_lhloose|					& \checkmark & - & - & - \\
\verb|HLT_e26_lhtight_nod0_ivarloose|	& - & \checkmark & \checkmark & \checkmark \\
\verb|HLT_e60_lhmedium_nod0|			& - & \checkmark & \checkmark & \checkmark \\
\verb|HLT_e140_lhloose_nod0|			& - & \checkmark & \checkmark & \checkmark \\
\midrule
\multicolumn{5}{c}{Di-electron triggers} \\
\midrule
\verb|HLT_2e12_lhloose_L12EM10VH |		& \checkmark & - & - & - \\
\verb|HLT_2e17_lhvloose_nod0| 			& - & \checkmark & - & - \\
\verb|HLT_2e24_lhvloose_nod0| 			& - & - & \checkmark & \checkmark \\
\verb|HLT_2e17_lhvloose_nod0_L12EM15VHI| 	& - & - & - & \checkmark \\
\midrule
\multicolumn{5}{c}{Single muon trigger} \\
\midrule
\verb|HLT_mu20_iloose_L1MU15| 			& \checkmark & - & - & - \\
\verb|HLT_mu40|							& \checkmark & - & - & - \\
\verb|HLT_mu26_ivarmedium| 				& - & \checkmark & \checkmark & \checkmark \\
\verb|HLT_mu50|							& - & \checkmark & \checkmark & \checkmark \\
\bottomrule\bottomrule
\end{tabular}
\end{table}

\section{Monte Carlo samples}
\label{sec:montecarlo}
Monte Carlo simulated samples are used to estimate signal acceptance before unblinding, profile the physics background for the analysis and to study object optimizations. Simulated samples for this analysis use are generated from ATLAS generalized MC20a/d/e samples for Run 2, using full detector simulation (\acs{FS}) and fast simulation (\acs{AF3}) to simulate detector response. \acs{MC} samples used and simulation processes are summarized in \autoref{tab:MC}.

\begin{table}[!htb]
\small
\setlength{\tabcolsep}{3pt} % Reduce column spacing
\renewcommand{\arraystretch}{1.2} % Adjust row spacing
\centering
\caption{\label{tab:MC}Summary of all Monte-Carlo samples used in this analysis. $V$ refers to an \acs{EW} ($W^\pm$/$Z$/$\gamma^*$) or Higgs boson. Matrix element (\acs{ME}) order refers to the order in \acs{QCD} of the perturbative calculation. Tune refers to the underlying-event tune of the parton shower (\acs{PS}) generator.}
\resizebox{\textwidth}{!}{%
\begin{tabular}{p{2.8cm} p{4.5cm} p{3.0cm} p{4.3cm} p{1.5cm} p{1.7cm} p{1.2cm}}
\toprule 
Process    & \acs{ME} Generator  & \acs{ME} Order & \acs{ME} \acs{PDF}  & \acs{PS}  & Tune & Sim. \\
\midrule \bottomrule
\multicolumn{7}{l}{\textbf{Signals}} \\ \midrule
\ttZp	   & \mgamc			& LO       & \nnpdfonelo    & \pythia & A14  & FS \\
\midrule \bottomrule
\multicolumn{7}{l}{\textbf{\tttt and \ttt}} \\ \midrule
\tttt      & \mgamc         & NLO      & \nnpdfnlo     & \pythia & A14  & AF3 \\
           & \mgamc         & NLO      & \mmhtlo       & \herwigseven & H7-UE-MMHT & AF3 \\
           & \sherpa        & NLO      & \nnpdfnnlo	   & \herwigseven & \sherpa  & FS \\
\ttt       & \mgamc         & LO       & \nnpdftwo     & \pythia & A14  & AF3 \\
\midrule \bottomrule
\multicolumn{7}{l}{\textbf{\ttV}} \\ \midrule
\ttH       & \powhegbox~v2  & NLO      & \nnpdfnlo 	   & \pythia & A14  & FS \\
           & \powhegbox~v2  & NLO      & \nnpdfnlo     & \herwigseven & H7.2-Default & FS \\
$\ttbar(Z/\gamma^*)$
		   & \mgamc         & NLO      & \nnpdfnlo     & \pythia & A14  & FS \\
           & \sherpa        & NLO      & \nnpdfnnlo    & \sherpa & \sherpa  & FS \\
\ttW       & \sherpa        & NLO      & \nnpdfnnlo    & \sherpa & \sherpa  & FS \\
           & \sherpa        & LO       & \nnpdfnnlo    & \sherpa & \sherpa  & FS \\
\midrule \bottomrule
\multicolumn{7}{l}{\textbf{\ttbar and Single-Top}} \\ \midrule
\ttbar     & \powhegbox~v2  & NLO      & \nnpdfnlo     & \pythia & A14  & FS \\
$tW$       & \powhegbox~v2  & NLO      & \nnpdfnlo     & \pythia & A14  & FS \\
$t(q)b$    & \powhegbox~v2  & NLO      & \nnpdfnlo (s) & \pythia & A14  & FS \\
           &                &          & \nnpdfnlo4f (t) &        &      & FS \\
$tWZ$      & \mgamc         & NLO      & \nnpdfnlo     & \pythia & A14  & FS \\
\tZ        & \mgamc         & LO       & \nnpdfnlo4f   & \pythia & A14  & FS \\
\midrule \bottomrule
\multicolumn{7}{l}{\textbf{\ttVV}} \\ \midrule
\ttWW      & \mgamc			& LO       & \nnpdfnlo     & \pythia & A14  & FS \\
\ttWZ      & \mg            & LO       & \nnpdfnlo     & \pythia & A14  & AF3 \\
\ttHH      & \mg            & LO       & \nnpdfnlo     & \pythia & A14  & AF3 \\
\ttWH      & \mg            & LO       & \nnpdfnlo     & \pythia & A14  & AF3 \\
\ttZZ      & \mg            & LO       & \nnpdfnlo     & \pythia & A14  & AF3 \\
\midrule \bottomrule
\multicolumn{7}{l}{\textbf{$V(VV)$+jets and $VH$}} \\ \midrule
$V$+jets   & \sherpa        & NLO      & \nnpdfnnlo    & \sherpa & \sherpa & FS \\
$VV$+jets  & \sherpa        & NLO      & \nnpdfnnlo    & \sherpa & \sherpa & FS \\
           &                & LO ($gg\to VV$) &        &         &         & FS \\
$VVV$+jets & \sherpa        & NLO      & \nnpdfnnlo	   & \sherpa & \sherpa & FS \\
$VH$       & \powhegbox~v2  & NLO      & \nnpdfaznlo   & \pythia & A14     & FS \\
\midrule \bottomrule
\end{tabular}%
} % End resizebox
\end{table}

\subsection{\ttZp signal samples}
Signal \ttZp samples were generated based on the simplified top-philic resonance model in section \ref{sec:ttZp}. Six $Z'$ mass points were utilized for the generation of the signal sample: 1000, 1250, 1500, 2000, 2500 and 3000 GeV. The top-$Z'$ coupling $c_t$ is chosen to be 1 for a narrow resonance peak, and the chirality angle $\theta$ is chosen to be $\pi/4$ to suppress loop production of $Z'$. The samples were then generated with \mgamc v.3.5.0 \citep{Alwall:2014hca} at \acs{LO} with the \nnpdfonelo \citep{Ball:2014uwa} \acs{PDF} set interfaced with \pythia \citep{Sjostrand:2014zea} using A14 tune and \nnpdftwo \acs{PDF} set for parton showering and hadronization. The resonance width is calculated to be $4\%$ for $c_t=1$.
%\textcolor{red}{plots: \HT, nJets, parameter comparison, interference, $m_{\ttbar}$ invariant mass}

\subsection{Background samples}

\subsubsection*{\acs{SM} \tttt background}
The nominal \acs{SM} \tttt sample was generated with \mgamc \citep{Alwall:2014hca} at \acs{NLO} in \acs{QCD} with the \nnpdfnlo \citep{Ball:2014uwa} \acs{PDF} set and interfaced with \pythia.230 \citep{Sjostrand:2014zea} using A14 tune \citep{ATL-PHYS-PUB-2014-021}. Decays for top quarks are simulated at \acs{LO} with \madspin \citep{Frixione:2007zp, Artoisenet:2012st} to preserve spin information, while decays for $b$- and $c$-hadrons are simulated with \evtgen v1.6.0 \citep{Lange:2001uf}. The renormalization and factorization scales $\mu_R$ and $\mu_F$ are set to $1/4\sqrt{m^2+\pT^2}$, which represents the sum of transverse mass of all particles generated from the \acs{ME} calculation \citep{Frederix:2017wme}. The ATLAS detector response was simulated with \acs{AF3}. Additional auxiliary \tttt samples are also generated to evaluate the impact of generator and \acs{PS} uncertainties as shown in \ref{tab:MC}.
\subsubsection*{\ttW background}
Nominal \ttW sample was generated using \sherpa v2.2.10 \citep{Bothmann:2019yzt} at \acs{NLO} in \acs{QCD} with the \nnpdfnnlo \citep{Ball:2014uwa} \acs{PDF} with up to one extra parton at \acs{NLO} and two at \acs{LO}, which are matched and merged with the \sherpa \acs{PS} based on Catani-Seymour dipole factorization \citep{Schumann:2007mg} using the MEPS@NLO prescription \citep{Hoeche:2011fd, Hoeche:2012yf, Catani:2001cc, Hoeche:2009rj} and a merging scale of 30 GeV. Higher-order \acs{ME} corrections are provided in \acs{QCD} by the OpenLoops 2 library \citep{Cascioli:2011va, Denner:2016kdg, Buccioni:2019sur} and in \acs{EW} from $\mathcal{O}(\alpha^3)+\mathcal{O}(\alpha^2_S\alpha^2)$ (LO3 \& NLO2) via two sets of internal event weights. An alternative sample with only \acs{EW} corrections at \acs{LO} from $\mathcal{O}(\alpha_S\alpha^3)$ (NLO3) diagrams were also simulated with the same settings.
\subsubsection*{$\ttbar(Z/\gamma^*)$ background}
Nominal $\ttbar(Z/\gamma^*)$ samples were generated separately for different ranges of dilepton invariant mass \acs{mll} to account for on-shell and off-shell $Z/\gamma^*$ production. Sample for \acs{mll} between 1 and 5 GeV was produced using \mgamc \citep{Alwall:2014hca} at \acs{NLO} with the \nnpdfnlo \citep{Ball:2014uwa} \acs{PDF} set, interfaced with \pythia.230 \citep{Sjostrand:2014zea} using A14 tune \citep{ATL-PHYS-PUB-2014-021} and \nnpdftwo \acs{PDF} set. Sample for $\acs{mll} < 5$ GeV was produced with \sherpa v2.2.10 \citep{Bothmann:2019yzt} at \acs{NLO} using \nnpdfnnlo \acs{PDF} set. To account for generator uncertainty, an alternative $\acs{mll}>5$ GeV sample was generated with identical settings to the low $m_{\ell\ell}$ sample. The ATLAS detector response was simulated with full detector simulation (\acs{FS}).
\subsubsection*{\ttH background}
Generation of \ttH background was done using \powhegbox \citep{Frixione:2007nw, Nason:2004rx,Frixione:2007vw, Alioli:2010xd} at \acs{NLO} in \acs{QCD} with the \nnpdfnlo \acs{PDF} \citep{Ball:2014uwa} set. The nominal sample is interfaced with \pythia.230 \citep{Sjostrand:2014zea} using the A14 tune \citep{ATL-PHYS-PUB-2014-021} and the \nnpdftwo \citep{Ball:2012cx} \acs{PDF} set. Detector response is simulated using \acs{FS}. An alternative \ttH sample generated similarly, but instead interfaced with \herwigseven.2.3 \citep{Bahr:2008pv, Bellm:2015jjp} to study the impact of parton shower and hadronization model. Detector response for the alternative sample is simulated using \acs{AF3}.
\subsubsection*{\ttt background}
The \ttt sample is generated using \mgamc \citep{Alwall:2014hca} at \acs{LO} in \acs{QCD}, interfaced with \pythia \citep{Sjostrand:2014zea} using the A14 tune \citep{ATL-PHYS-PUB-2014-021}. The sample is produced in the five-flavor scheme \citep{Alekhin_2010} to prevent \acs{LO} interference with \tttt.
\subsubsection*{\ttbar background}
The \ttbar sample is modeled with \powhegbox \citep{Frixione:2007nw,Nason:2004rx,Frixione:2007vw,Alioli:2010xd} at \acs{NLO} in \acs{QCD} with the \nnpdfnlo \citep{Ball:2014uwa} \acs{PDF} set and the $h_\text{damp}$ parameter set to $1.5m_\text{top}$ \citep{ATL-PHYS-PUB-2016-020}. Events are interfaced with \pythia.230 \citep{Sjostrand:2014zea} using the A14 tune \citep{ATL-PHYS-PUB-2014-021} and the \nnpdftwo \citep{Ball:2012cx} \acs{PDF} set.
\subsubsection*{Single-top ($tW$ \& $t(q)b$) background}
Single-top $tW$-associated production is modeled using the \powhegbox generator \citep{Frixione:2007nw,Nason:2004rx,Frixione:2007vw,Alioli:2010xd} at \acs{NLO} in \acs{QCD} in the five-flavor scheme \citep{Alekhin_2010} with the \nnpdfnlo \citep{Ball:2014uwa} \acs{PDF} set. Interference with \ttbar production \citep{ATL-PHYS-PUB-2016-020} is handled using the diagram removal scheme \citep{Frixione:2008yi}. Single-top $t(q)b$ production is modeled using the \powhegbox generator at \acs{NLO} in \acs{QCD} with the s-channel production modeled in the five-flavor scheme with the \nnpdfnlo \acs{PDF} set, while the t-channel production is modeled in the four-flavor scheme with the \nnpdfnlo 4f \citep{Ball:2014uwa} \acs{PDF} set. The \ttWW contributions are normalized to \acs{NLO} theoretical cross section. All single-top samples are interfaced with \pythia.230 \citep{Sjostrand:2014zea} using the A14 tune \citep{ATL-PHYS-PUB-2014-021} and the \nnpdftwo \citep{Ball:2012cx} \acs{PDF} set.
\subsubsection*{\tWZ +jets background}
The \tWZ sample is generated using \mgamc \citep{Alwall:2014hca} at \acs{NLO} in \acs{QCD} with the \nnpdfnlo \citep{Ball:2014uwa} \acs{PDF} set, interfaced with \pythia.212 \citep{Sjostrand:2014zea} using the A14 tune \citep{ATL-PHYS-PUB-2014-021} and the \nnpdftwo \citep{Ball:2012cx} \acs{PDF} set.
\subsubsection*{$tZ$ \& \ttVV background}
Production of $tZ$ is modeled using \mgamc \citep{Alwall:2014hca} at \acs{NLO} in \acs{QCD} with scale of $\HT/6$ and the \nnpdfnlo 4f \citep{Ball:2014uwa} \acs{PDF} set. Production of \ttWW is modeled using \mgamc \citep{Alwall:2014hca} at \acs{LO}, while production of $\ttbar WZ$, $\ttbar HH$, $\ttbar WH$ and $\ttbar ZZ$ are modeled using \mg at \acs{LO}. All \ttVV samples use the \nnpdfnlo \citep{Ball:2014uwa} \acs{PDF} set, and all samples in this section are interfaced with \pythia \citep{Sjostrand:2014zea} using the A14 tune \citep{ATL-PHYS-PUB-2014-021}.
\subsubsection*{Single boson ($V$) +jets background}
Production of $V$+jets is modeled with \sherpa v2.2.10 \citep{Bothmann:2019yzt} using \acs{NLO} \acs{ME} for up to two jets and \acs{LO} \acs{ME} for up to four jets, with the \nnpdfnlo \citep{Ball:2014uwa} \acs{PDF} set. Matrix elements are calculated with the Comix \citep{Gleisberg:2008fv} and OpenLoops libraries \citep{Cascioli:2011va, Denner:2016kdg} and matched with the \sherpa \acs{PS} based on Catani-Seymour dipole factorization \citep{Schumann:2007mg} using the MEPS@NLO prescription \citep{Hoeche:2011fd, Hoeche:2012yf, Catani:2001cc, Hoeche:2009rj}. The sample is normalized to \acs{NNLO} \citep{Anastasiou:2003ds} theoretical cross section.
\subsubsection*{Diboson ($VV$) +jets background}
Diboson samples are simulated with \sherpa v2.2.14 \citep{Bothmann:2019yzt} with the \nnpdfnlo \citep{Ball:2014uwa} \acs{PDF} set. Fully leptonic and semileptonic final states are generated using \acs{NLO} \acs{ME} for up to one extra parton and  \acs{LO} \acs{ME} for up to three extra parton emissions. Loop-induced processes are generated using \acs{LO} \acs{ME} for up to one extra parton. Matrix elements are matched and merged with the \sherpa \acs{PS} based on Catani-Seymour dipole factorization \citep{Schumann:2007mg} using the MEPS@NLO prescription \citep{Hoeche:2011fd, Hoeche:2012yf, Catani:2001cc, Hoeche:2009rj}. Virtual \acs{QCD} \acs{ME} corrections are provided by the OpenLoops library \citep{Cascioli:2011va, Denner:2016kdg}.
\subsubsection*{Triboson ($VVV$) +jets background}
The triboson sample is modeled with \sherpa v2.2.10 \citep{Bothmann:2019yzt} using factorized gauge boson decays. Matrix elements for the inclusive process at \acs{NLO} and up to two extra partons at \acs{LO} are matched and merged with the \sherpa \acs{PS} based on Catani-Seymour dipole factorization \citep{Schumann:2007mg} using the MEPS@NLO prescription \citep{Hoeche:2011fd, Hoeche:2012yf, Catani:2001cc, Hoeche:2009rj}. Virtual \acs{QCD} \acs{ME} corrections are provided by the OpenLoops library \citep{Cascioli:2011va, Denner:2016kdg}.
\subsubsection*{$VH$ background}
Generation of $WH$ and $ZH$ samples is performed using \powhegbox \citep{Frixione:2007nw,Nason:2004rx,Frixione:2007vw,Alioli:2010xd} at \acs{NLO} with the \nnpdfaznlo \citep{Ball:2014uwa} \acs{PDF} set, interfaced with \pythia.230 \citep{Sjostrand:2014zea} using the A14 tune \citep{ATL-PHYS-PUB-2014-021} and the \nnpdftwo \citep{Ball:2012cx} \acs{PDF} set. The samples are normalized to theoretical cross sections at \acs{NNLO} in \acs{QCD} and \acs{NLO} in \acs{EW} accuracies.

\end{document}